Short-time expansion of Heisenberg operators in open collective quantum spin systems

### 2019-12-12 soc_squeezing/collective_paper/review/response.tex

  \red{Our method is based on a short-time expansion of exact
    solutions to the equations of motion for Heisenberg operators.
    Evaluating this expansion requires ...}

  to:

  \green{Our method is based on a formal solution to the equations of
    motion for Heisenberg operators, thereby bearing some resemblance
    to the Mori formalism[59] and related work[60].  Specifically, we
    expand a formal solution for a Heisenberg operator into a Taylor
    series whose truncation can yield negligible error at sufficiently
    short times.  Evaluating the resulting expansion requires ...}

  We have also added a reference to these works to the discussion of
  the limitations of the TST expansion at the end of the theory
  section of our manuscript (right column of page 3), changing the
  text:

  \red{Precisely characterizing the implications of these last two
    considerations for the TST expansion requires additional analysis
    that we defer to future work.  As we show from benchmarks of the
    TST expansion in Section III, however, a detailed understanding of
    these limitations is not necessary ...}

@@

  to:

  \green{Precisely characterizing the implications of these last two
    considerations for the TST expansion requires additional analysis
    that we defer to future work.  {\bf An investigation of
      connections between the TST expansion and past work related to
      the Mori formalism[59, 60], for example, might answer questions
      about the breakdown and convergence of the TST expansion.}  As
    we show from benchmarks of the TST expansion in Section III,
    however, a detailed understanding of {\bf breakdown} is not
    necessary ...}

@@

\item Concerning the second question, about decoherence:

  \blue{2. I am not sure if the paper properly takes into account
    previous results about the influence of decoherence on spin
    squeezing. First, there exists many mechanisms for decoherence,
    while decoherence considered in this paper is due to uncorrelated
    decay of individual spins. This is not clarified in the
    introduction. In fact, atom loss represents an important mechanism
    of decoherence in spin squeezing experiments, but is not
    considered in this paper. Could the method studied here also be
    applied to atom loss? It would be helpful if this point is
    mentioned. Second, for the specific decoherence model considered,
    the authors only cite a few studies about the influence of such
    decoherence on spin squeezing from one-axis twisting (OAT). There
    are many studies about the effect of decoherence on spin
    squeezing.}

  The referee makes a fair point that we claim to present a method for
  simulating ``generic'' Markovian decoherence, when in fact we only
  address certain (typical) forms of Markovian decoherence.  To avoid
  misleading readers about the contents of our paper, we have
  therefore replaced the words ``\red{generic Markovian decoherence}''
  by ``\green{typical forms of Markovian decoherence}'' in the
  abstract, as well as the last paragraph of the introduction.

  We would like to point out that our work {\it does} address more
  than just uncorrelated decay of individual spins: we also address
  uncorrelated excitation and dephasing of individual spins, as well
  as arbitrary coherent or incoherent combinations of these processes.
  Similarly, we have done all of the work necessary to account for
  arbitrary combinations of {\it collective} decay, excitation, and
  dephasing in Appendices D-G, as mentioned in the paragraph below
  Eq.~(4) of our manuscript.

@@

  The referee is correct, though, that our original manuscript does
  not discuss atom loss, which is an important decoherence mechanism
  in some experimental realizations of collective spin models.  We
  have therefore added the following discussion about atom loss to the
  theory section our manuscript (bottom right of page 2):

  \green{We note that particle loss is an important decoherence
    mechanism in many experimental realizations of collective spin
    models[41].  In principle, a spin model has no notion of the
    particle annihilation operators that generate particle loss, and
    therefore cannot capture this effect directly.  Nonetheless, for a
    system initially composed of $N$ particles, the effect of particle
    loss can be emulated with $O(1/N)$ error by the dissipator
    $\check\D_{\t{loss}}$ defined by
    $\check\D_{\t{loss}}\hat\S_{\v m}=-\abs{\v m}\hat\S_{\v m}$, where
    $\abs{\v m}\equiv\sum_\alpha m_\alpha$ (see Appendix H).
    Furthermore, the effect particle loss can be accounted for exactly
    by (i) introducing an additional index on spin operators,
    $\hat\S_{\v m}\to\hat\S_{N\v m}$, to keep track of different
    sectors of fixed particle number within a multi-particle Fock
    space, and (ii) constructing jump operators that appropriately
    couple spin operators within different particle-number sectors.
    We defer a detailed exact accounting of particle loss to future
    work.}

  The new appendix provides mathematical justification for our claim
  that the dissipator $\check\D_{\t{loss}}$ emulates particle loss
  with $O(1/N)$ error.  We hope that these additions do a satisfactory
  job of addressing particle loss.

  As a last point, concerning the number of references in our work to
  past studies about the influence of decoherence on spin squeezing in
  the one-axis twisting (OAT) model: the small number of such
  references is due to the fact that we do not have any particular
  discussions of the influence of decoherence on spin squeezing with
  OAT, and do not consider such discussions relevant to our current
  work.  Even with decoherence (of the types primarily considered in
  our work), the OAT model is an exactly analytically solvable (see
  \url{https://doi.org/gcsjdp}).  Approximate numerical methods such
  as that presented in our manuscript are therefore inappropriate
  tools for the study of OAT.  The primary role of OAT in our work is
  to serve as a simple and exactly solvable benchmark of our numerical
  methods, rather than an object of study in and of itself.  To
  clarify this point, we have changed the second paragraph of section
  III on spin squeezing, benchmarking, and breakdown (top of page 4),
  from:

@@

  \red{Note that the OAT model is a special case of the zero-field
    Ising model, whose quantum dynamics admits an exact analytic
    solution even in the presence of decoherence[62].  We will
    benchmark our calculations using these analytical OAT results
    wherever applicable (see Appendix J, as well as the Supplementary
    Material of Ref.~[14]).}

  to:

  \green{Note that the OAT model is a special case of the zero-field
    Ising model, whose quantum dynamics admits an exact analytic
    solution even in the presence of decoherence[64].  The approximate
    and numerics-oriented TST expansion is therefore an inappropriate
    tool for studying the OAT model, which will merely serve as an
    exactly solvable benchmark of our methods.  Wherever applicable,
    we will provide exact results for the OAT model (see Appendix K,
    as well as the Supplementary Material of Ref.~[14]).}

\item Concerning the final question, about the advantages of the TNT
  model in the presence of decoherence:

@@

  \blue{3. I am surprised by the results from the strong decoherence
    regimes in Fig.~2, which shows that twist and-turn (TNT) model is
    much more robust than OAT and TAT. Is this physical or just an
    artifact from numerical errors. To rule out the latter, I suggest
    the authors check the results in the strong decoherence regime by
    studying a smaller system, where Monte Carlo simulation is
    accessible and can be used as a comparison. If this result is
    indeed physical, if would be very helpful if the authors could
    comment on the reason?}

  We thank the referee for their careful attention to our results.  We
  agree that our findings warrant a comment on why TNT performs better
  than OAT or TAT in the presence of decoherence.  We have therefore
  rewritten most of the last paragraph of section III, changing:

  \red{The results in Figure 2 show that TNT is much more robust to
    single-spin decoherence than OAT or TAT, a finding that could not
    necessarily be deduced from the weak-decoherence simulations
    presented in Figure 1.  Strong-decoherence computations of this
    sort ...}

  \green{The results in Figure 2 show that the TNT model can generate
    more squeezing than the OAT or TAT models in the presence of
    strong decoherence.  The better performance of TNT is a
    consequence of the fact that TNT initially generates squeezing at
    a faster rate than OAT or TAT, thereby allowing it to produce more
    squeezing before the degrading effects of decoherence kick in.  We
    corroborate this finding with quantum trajectory simulations of a
    small system in Appendix L.  Strong-decoherence computations of
    the sort used for Figure 2 ...}

  The new appendix includes Monte Carlo simulations of a small system
  with sufficiently strong decoherence to reproduce the improved
  performance of TNT.  We hope that these additions address the
  referee's questions and concerns about figure 2 to a satisfactory
  degree.

  % One consequence of applying our method in the large system, strong
  % decoherence regime in figure 2 is that there is no way (barring a
  % detailed analysis of error and breakdown in the TST expansion) to
  % completely rule out numerical errors, as there are no other
  % simulation methods available to corroborate the results.  We had
  % hoped that the benchmarking in figure 1 and the discussion of
  % breakdown mechanisms would assure readers that the TST expansion
  % self-diagnoses its sudden departure from virtually exact results,
  % thereby increasing confidence that the TST expansion can be
  % trusted when no other methods are available.

@@

  \red{The growth of genuinely multi-body correlations, however,
    eventually causes the cumulant expansion to yield unphysical
    results with no clear signature of failure.}

  to:

  \green{The growth of genuinely multi-body correlations, however,
    eventually causes the cumulant expansion to yield incorrect
    results with no clear signature of failure.}

\item One of our previous changes to the manuscript defined
  ``\green{$\abs{\v m}\equiv\sum_\alpha m_\alpha$}'' (bottom of page 2
  in the new text).  We have therefore removed a subsequent definition
  of this notation, ``\red{where
    $\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$}'', which previously
  appeared in the following paragraph (middle right column of page 2
  in the old text).

### 2019-12-12 soc_squeezing/collective_paper/collective_spin.tex
penultimate (I hope) changes in response to referee on collective spin paper

\begin{abstract}
  We present a new method to compute short-time expectation values in
  large collective spin systems with generic Markovian decoherence.
  Our method is based on a Taylor expansion of a formal solution to
  the equations of motion for Heisenberg operators.  This expansion
  can be truncated at finite order to obtain virtually exact results
  at short times that are relevant for metrological applications such
  as spin squeezing.  In order to evaluate the expansion for
  Heisenberg operators, we compute the relevant structure constants of
  a collective spin operator algebra.  We demonstrate the utility of
  our method by computing spin squeezing, two-time correlation
  functions, and out-of-time-ordered correlators for $10^4$ spins in
  strong-decoherence regimes that are otherwise inaccessible via
  existing numerical methods.  Our method can be straightforwardly
  generalized to the case of a collective spin coupled to bosonic
  modes, relevant for trapped ion and cavity QED experiments, and may
  be used to investigate short-time signatures of quantum chaos and
  information scrambling.
\end{abstract}

@@

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superselection sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence generally violates total spin conservation and
requires the use of density operators, increasing the dimension of
accessible state space to $O\p{N^3}$\cite{hartmann2016generalized,
  xu2013simulating}.  In this case, exact simulations can be carried
out for $N\lesssim100$ particles.  If decoherence is sufficiently
weak, dynamics can be numerically solvable for $N\lesssim10^5$
particles via ``quantum trajectory'' Monte Carlo
methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
as ``quantum jump'' or ``Monte Carlo wavefunction'' methods) that can
reproduce all expectation values of interest.  When decoherence is
strong, however, these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct quantum trajectories
that need to be averaged in order to accurately compute expectation
values.  Even with strong decoherence, dynamics are sometimes solvable
through the cumulant expansion\cite{meiser2010steadystate} that
neglects all $n$-body connected correlators for $n>2$.  The growth of
genuinely multi-body correlations, however, eventually causes the
cumulant expansion to yield unphysical results with no clear signature
of failure.  In the absence of other means to compute correlators, it
can therefore be difficult to identify the point at which correlators
computed via the cumulant expansion can no longer be trusted.

In this work, we present a new method to compute short-time dynamics
of collective spin systems with generic Markovian decoherence.  The
only restriction on decoherence (beyond Markovianity) is that, like
the coherent collective dynamics, it must act identically on all
constituent particles.  Our method is based on a short-time expansion
of exact solutions to the equations of motion for Heisenberg
operators.  Evaluating this expansion requires knowing the structure
constants of a collective spin operator algebra; the calculation of
these structure constants (in Appendices
\ref{sec:identities}--\ref{sec:general_product}) is one of the main
technical results of this work, which we hope will empower both
analytical and numerical studies of collective spin systems in the
future.  We benchmark our method against exact results from both
analytical calculations and quantum trajectory Monte Carlo
computations of spin squeezing in accessible parameter regimes,
highlighting both advantages and limitations of the short-time
expansion.  Finally, we showcase applications of our method by
computing quantities that are inaccessible to other numerical methods.

@@

The time derivative operator $\check T$ will generally couple spin
operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
signifies the growth of many-body correlations.  Keeping track of this
growth eventually becomes intractable, requiring us to truncate our
equations of motion somehow.  The simplest truncation strategy would
be to take
\begin{align}
  \f{d}{dt} \bk{\hat\O\p{t}}
  \to \sum_{w\p{\v m}<W} \bk{\hat\S_{\v m}}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
high-weight cutoff $W$.  The truncation in
Eq.~\eqref{eq:weight_truncation} closes the system of differential
equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
using standard numerical methods.  Some initial conditions for this
system of differential equations, namely expectation values of
collective spin operators with respect to spin-polarized (Gaussian)
states that are generally simple to prepare experimentally, are
provided in Appendix \ref{sec:initial_conditions}.

@@

Three primary considerations limit the maximum time $t$ to which we
can accurately compute a correlator $\bk{\hat S_{\v n}\p{t}}$ using
the TST expansion.  First, maintaining accuracy at larger times $t$
requires going to higher orders $M$ in the TST expansion.  An
order-$M$ TST expansion of the correlator $\bk{\hat S_{\v n}\p{t}}$
can involve a significant fraction of operators $\hat S_{\v m}$ with
weight $\abs{\v m}\lesssim M$, which implies the need to compute
$O\p{M^3}$ initial-time expectation values $\bk{\hat S_{\v m}}$ and
$O\p{M^4}$ matrix elements $T^k_{\v m\v n}$.  In practice, with a
straightforward implementation of the TST expansion we find that these
requirements generally restrict $M\lesssim 50$ -- $70$ with $8$ --
$50$ gigabytes of random access memory (RAM).  Second, individual
terms at high orders of the TST expansion in Eq.~\eqref{eq:TST} can
grow excessively large, greatly amplifying any numerical errors and
thereby spoiling cancellations that are necessary to arrive at a
physical value of a correlator, i.e.~with
$\abs{\bk{\hat S_{\v n}\p{t}}}\lesssim S^{\abs{\v n}}$.  Finally, the
TST expansion is essentially perturbative in the time $t$, which
implies that its validity as a formal expansion eventually breaks
down.  Precisely characterizing the implications of these last two
considerations for the TST expansion requires additional analysis that
we defer to future work.  As we show from benchmarks of the TST
expansion in Section \ref{sec:squeezing}, however, a detailed
understanding of these limitations is not necessary to diagnose the
breakdown time $t_{\t{break}}^{(M)}$ beyond which the TST expansion
yields inaccurate results.  Empirically, we find that going beyond
order $M\approx35$ yields no significant gains in all cases considered
in this work.

@@

Note that the OAT model is a special case of the zero-field Ising
model, whose quantum dynamics admits an exact analytic solution even
in the presence of decoherence\cite{foss-feig2013nonequilibrium}.  We
will benchmark our calculations using these analytical OAT results
wherever applicable (see Appendix \ref{sec:OAT}, as well as the
Supplementary Material of Ref.~[\citenum{bohnet2016quantum}]).

@@

The main lesson from Figure \ref{fig:benchmarking} is that the TST
expansion yields essentially exact results right up until a sudden and
drastic departure that can be diagnosed by inspection.  The breakdown
of the TST expansion in Figure \ref{fig:benchmarking} is marked by an
unphysical squeezing parameter $\xi^2<0$, which occurs due to
individual correlators taking unphysical values with
$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  Going up
through order $M=70$ in the TST expansion does not significantly
increase the breakdown time $t_{\t{break}}^{(M)}$, and in some cases
even shortens $t_{\t{break}}^{(M)}$.  The sudden and drastic departure
from virtually exact results is consistent with the limitations of the
TST expansion discussed at the end of Section \ref{sec:theory}.
Specifically, we identify three possible mechanisms for breakdown:
\begin{enumerate*}
\item a rapid growth in the order $M$ necessary for the TST expansion
  to converge,
\item the growth of numerical errors in excessively large terms of the
  TST expansion, and
\item the formal breakdown of the perturbative expansion in the time
  $t$.
\end{enumerate*}
In all of these cases, a detailed cancellation eventually ceases to
occur between large terms at high orders in the TST expansion.  These
large terms grow with the time $t$ raised to some large power (as high
as $M$), and therefore rapidly yield wildly unphysical results.  In
contrast to other approximate methods such as the cumulant
expansion\cite{meiser2010steadystate}, the TST expansion can thus
diagnose its own breakdown, which is an important feature when working
in parameter regimes that are inaccessible via other means to compute
correlators.

@@

The results in Figure \ref{fig:decoherence_strong} show that TNT is
much more robust to single-spin decoherence than OAT or TAT, a finding
that could not necessarily be deduced from the weak-decoherence
simulations presented in Figure \ref{fig:benchmarking}.
Strong-decoherence computations of this sort were used to put lower
bounds on theoretically achievable spin squeezing via TAT with
decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete and practical application of the TST expansion and the
collective-spin structure constants calculated in this work.

@@

We have presented a new method for computing correlators at short
times in collective spin systems.  This method is based on truncating
a short-time expansion of Heisenberg operators, and can access
correlators on time scales that are relevant to metrological
applications such as spin squeezing.  In order to evaluate the
truncated short-time (TST) expansion of Heisenberg operators, we have
computed the structure constants of a collective spin operator
algebra, which we hope will empower future analytical and numerical
studies of collective spin systems.  Even though we considered only
non-equilibrium spin-squeezing processes in this work, our method can
be applied directly in an equilibrium setting, and is straightforward
to generalize to systems such as trapped ions and optical cavities
with collective spin-boson interactions.  In such contexts, our method
may be used to benchmark the short-time effects of decoherence, or
study the onset of quantum chaos and information scrambling.

@@

@@

>>>>

\begin{abstract}
  We present an efficient method to compute short-time expectation
  values in large collective spin systems with typical forms of
  Markovian decoherence.  Our method is based on a Taylor expansion of
  a formal solution to the equations of motion for Heisenberg
  operators.  This expansion can be truncated at finite order to
  obtain virtually exact results at short times that are relevant for
  metrological applications such as spin squeezing.  In order to
  evaluate the expansion for Heisenberg operators, we compute the
  relevant structure constants of a collective spin operator algebra.
  We demonstrate the utility of our method by computing spin
  squeezing, two-time correlation functions, and out-of-time-ordered
  correlators for $10^4$ spins in strong-decoherence regimes that are
  otherwise inaccessible via existing numerical methods.  Our method
  can be straightforwardly generalized to the case of a collective
  spin coupled to bosonic modes, relevant for trapped ion and cavity
  QED experiments, and may be used to investigate short-time
  signatures of quantum chaos and information scrambling.
\end{abstract}

@@

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superselection sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence generally violates total spin conservation and
requires the use of density operators, increasing the dimension of
accessible state space to $O\p{N^3}$\cite{hartmann2016generalized,
  xu2013simulating}.  In this case, exact simulations can be carried
out for $N\lesssim100$ particles.  If decoherence is sufficiently
weak, dynamics can be numerically solvable for $N\lesssim10^5$
particles via ``quantum trajectory'' Monte Carlo
methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
as ``quantum jump'' or ``Monte Carlo wavefunction'' methods) that can
reproduce all expectation values of interest.  When decoherence is
strong, however, these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct quantum trajectories
that need to be averaged in order to accurately compute expectation
values.  Even with strong decoherence, dynamics are sometimes solvable
through the cumulant expansion\cite{meiser2010steadystate} that
neglects all $n$-body connected correlators for $n>2$.  The growth of
genuinely multi-body correlations, however, eventually causes the
cumulant expansion to yield incorrect results with no clear signature
of failure.  In the absence of other means to compute correlators, it
can therefore be difficult to identify the point at which correlators
computed via the cumulant expansion can no longer be trusted.

In this work, we present an efficient method to compute short-time
dynamics of collective spin systems with typical forms of Markovian
decoherence.  The only restriction on decoherence (beyond
Markovianity) is that, like the coherent collective dynamics, it must
act identically on all constituent particles.  Our method is based on
a formal solution to the equations of motion for Heisenberg operators,
thereby bearing some resemblance to the Mori
formalism\cite{mori1965continuedfraction} and related
work\cite{annett1994recursive}.  Specifically, we expand a formal
solution for a Heisenberg operator into a Taylor series whose
truncation can yield negligible error at sufficiently short times.
Evaluating the resulting expansion requires knowing the structure
constants of a collective spin operator algebra; the calculation of
these structure constants (in Appendices
\ref{sec:identities}--\ref{sec:general_product}) is one of the main
technical results of this work, which we hope will empower both
analytical and numerical studies of collective spin systems in the
future.  We benchmark our method against exact results from both
analytical calculations and quantum trajectory Monte Carlo
computations of spin squeezing in accessible parameter regimes,
highlighting both advantages and limitations of the short-time
expansion.  Finally, we showcase applications of our method by
computing quantities that are inaccessible to other numerical methods.

@@

We note that particle loss is an important decoherence mechanism in
many experimental realizations of collective spin
models\cite{ma2011quantum}.  In principle, a spin model has no notion
of the particle annihilation operators that generate particle loss,
and therefore cannot capture this effect directly.  Nonetheless, for a
system initially composed of $N$ particles, the effect of particle
loss can be emulated with $O(1/N)$ error by the dissipator
$\check\D_{\t{loss}}$ defined by
$\check\D_{\t{loss}}\hat\S_{\v m}=-\abs{\v m}\hat\S_{\v m}$, where
$\abs{\v m}\equiv\sum_\alpha m_\alpha$ (see Appendix
\ref{sec:particle_loss}).  Furthermore, the effect particle loss can
be accounted for exactly by
\begin{enumerate*}
\item introducing an additional index on spin operators,
  $\hat\S_{\v m}\to\hat\S_{N\v m}$, to keep track of different sectors
  of fixed particle number within a multi-particle Fock space, and
\item constructing jump operators that appropriately couple spin
  operators within different particle-number sectors.
\end{enumerate*}
We defer a detailed exact accounting of particle loss to future work.

The time derivative operator $\check T$ will generally couple spin
operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$.  The growth of
operator weight signifies the growth of many-body correlations.
Keeping track of this growth eventually becomes intractable, requiring
us to truncate our equations of motion somehow.  The simplest
truncation strategy would be to take
\begin{align}
  \f{d}{dt} \bk{\hat\O\p{t}}
  \to \sum_{w\p{\v m}<W} \bk{\hat\S_{\v m}}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
high-weight cutoff $W$.  The truncation in
Eq.~\eqref{eq:weight_truncation} closes the system of differential
equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
using standard numerical methods.  Some initial conditions for this
system of differential equations, namely expectation values of
collective spin operators with respect to spin-polarized (Gaussian)
states that are generally simple to prepare experimentally, are
provided in Appendix \ref{sec:initial_conditions}.

@@

Three primary considerations limit the maximum time $t$ to which we
can accurately compute a correlator $\bk{\hat S_{\v n}\p{t}}$ using
the TST expansion.  First, maintaining accuracy at larger times $t$
requires going to higher orders $M$ in the TST expansion.  An
order-$M$ TST expansion of the correlator $\bk{\hat S_{\v n}\p{t}}$
can involve a significant fraction of operators $\hat S_{\v m}$ with
weight $\abs{\v m}\lesssim M$, which implies the need to compute
$O\p{M^3}$ initial-time expectation values $\bk{\hat S_{\v m}}$ and
$O\p{M^4}$ matrix elements $T^k_{\v m\v n}$.  In practice, with a
straightforward implementation of the TST expansion we find that these
requirements generally restrict $M\lesssim 50$ -- $70$ with $8$ --
$50$ gigabytes of random access memory (RAM).  Second, individual
terms at high orders of the TST expansion in Eq.~\eqref{eq:TST} can
grow excessively large, greatly amplifying any numerical errors and
thereby spoiling cancellations that are necessary to arrive at a
physical value of a correlator, i.e.~with
$\abs{\bk{\hat S_{\v n}\p{t}}}\lesssim S^{\abs{\v n}}$.  Finally, the
TST expansion is essentially perturbative in the time $t$, which
implies that its validity as a formal expansion eventually breaks
down.  Precisely characterizing the implications of these last two
considerations for the TST expansion requires additional analysis that
we defer to future work.  An investigation of connections between the
TST expansion and past work related to the Mori
formalism\cite{mori1965continuedfraction, annett1994recursive}, for
example, might answer questions about the breakdown and convergence of
the TST expansion.  As we show from benchmarks of the TST expansion in
Section \ref{sec:squeezing}, however, a detailed understanding of
breakdown is not necessary to diagnose the breakdown time
$t_{\t{break}}^{(M)}$ beyond which the TST expansion yields inaccurate
results.  Empirically, we find that going beyond order $M\approx35$
yields no significant gains in all cases considered in this work.

@@

Note that the OAT model is a special case of the zero-field Ising
model, whose quantum dynamics admits an exact analytic solution even
in the presence of decoherence\cite{foss-feig2013nonequilibrium}.  The
approximate and numerics-oriented TST expansion is therefore an
inappropriate tool for studying the OAT model, which will merely serve
as an exactly solvable benchmark of our methods.  Wherever applicable,
we will provide exact results for the OAT model (see Appendix
\ref{sec:OAT}, as well as the Supplementary Material of
Ref.~[\citenum{bohnet2016quantum}]).

@@

The main lesson from Figure \ref{fig:benchmarking} is that the TST
expansion yields essentially exact results right up until a sudden and
drastic departure that can be diagnosed by inspection.  The breakdown
of the TST expansion in Figure \ref{fig:benchmarking} induces an
unphysical squeezing parameter $\xi^2<0$.  In general, however, there
is no fundamental relationship between the breakdown of the TST
expansion and the conditions for a physical squeezing parameter
$\xi^2$.  A proper diagnosis of breakdown therefore requires
inspection of the correlators $\bk{\hat\S_{\v n}\p{t}}$ used to
compute the squeezing parameter $\xi^2$, which upon breakdown will
rapidly take unphysical values with
$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  The sudden and
drastic departure from virtually exact results is consistent with the
limitations of the TST expansion discussed at the end of Section
\ref{sec:theory}.  Specifically, we identify three possible mechanisms
for breakdown:
\begin{enumerate*}
\item a rapid growth in the order $M$ necessary for the TST expansion
  to converge,
\item the growth of numerical errors in excessively large terms of the
  TST expansion, and
\item the formal breakdown of the perturbative expansion in the time
  $t$.
\end{enumerate*}
In all of these cases, a detailed cancellation eventually ceases to
occur between large terms at high orders in the TST expansion.  These
large terms grow with the time $t$ raised to some large power (as high
as $M$), and therefore rapidly yield wildly unphysical results.  In
contrast to other approximate methods such as the cumulant
expansion\cite{meiser2010steadystate}, the TST expansion can thus
diagnose its own breakdown, which is an important feature when working
in parameter regimes that are inaccessible via other means to compute
correlators.  Note that, due to the breakdown mechanisms of the TST
expansion, going up through order $M=70$ does not significantly
increase the breakdown time $t_{\t{break}}^{(M)}$ in Figure
\ref{fig:benchmarking}, and in some cases even shortens
$t_{\t{break}}^{(M)}$.

@@

The results in Figure \ref{fig:decoherence_strong} show that the TNT
model can generate more squeezing than the OAT or TAT models in the
presence of strong decoherence.  The better performance of TNT is a
consequence of the fact that TNT initially generates squeezing at a
faster rate than OAT or TAT, thereby allowing it to produce more
squeezing before the degrading effects of decoherence kick in.  We
corroborate this finding with quantum trajectory simulations of a
small system in Appendix \ref{sec:small_system}.  Strong-decoherence
computations of the sort used for Figure \ref{fig:decoherence_strong}
were used to put lower bounds on theoretically achievable spin
squeezing via TAT with decoherence in
Ref.~[\citenum{he2019engineering}], exemplifying a concrete and
practical application of the TST expansion and the collective-spin
structure constants calculated in this work.

@@

We have presented an efficient method for computing correlators at
short times in collective spin systems.  This method is based on
truncating a short-time expansion of Heisenberg operators, and can
access correlators on time scales that are relevant to metrological
applications such as spin squeezing.  In order to evaluate the
truncated short-time (TST) expansion of Heisenberg operators, we have
computed the structure constants of a collective spin operator
algebra, which we hope will empower future analytical and numerical
studies of collective spin systems.  Even though we considered only
non-equilibrium spin-squeezing processes in this work, our method can
be applied directly in an equilibrium setting, and is straightforward
to generalize to systems such as trapped ions and optical cavities
with collective spin-boson interactions.  In such contexts, our method
may be used to benchmark the short-time effects of decoherence, or
study the onset of quantum chaos and information scrambling.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Emulating particle loss in a spin model}
\label{sec:particle_loss}

Here we discuss the details of emulating particle loss with $O(1/N)$
error, where $N$ is the initial number of particles in a system that
we wish to describe with a spin model.  Starting with the full algebra
of creation and annihilation operators (whether bosonic or fermionic)
in a system, spin models are typically implemented by identifying a
subalgebra of relevant ``spin'' operators that satisfy appropriate
commutation relations.  Two-state particles on a lattice, for example,
are described by annihilation operators $c_{js}$ indexed by a lattice
site $j\in\mathbb{Z}$ and an internal state index $s\in\set{\up,\dn}$,
enabling the straightforward construction of spin operators
\begin{align}
  \sigma_\x^{(j)} \equiv c_{j,\up}^\dag c_{j,\dn} + \t{h.c.},
  &&
  \sigma_\y^{(j)} \equiv - i c_{j,\up}^\dag c_{j,\dn} + \t{h.c.},
  &&
  \sigma_\z^{(j)}
  \equiv c_{j,\up}^\dag c_{j,\up} - c_{j,\dn}^\dag c_{j,\dn},
  &&
  \1^{(j)} \equiv c_{j,\up}^\dag c_{j,\up} + c_{j,\dn}^\dag c_{j,\dn},
\end{align}
which satisfy same commutation relations as the standard Pauli
operators.  These spin operators can be more compactly defined in the
form
\begin{align}
  \sigma_\alpha^{(j)} \equiv \sum_{r,s\in\set{\up,\dn}}
  c_{jr}^\dag \bk{r|\sigma_\alpha|s} c_{js},
\end{align}
where $\sigma_\alpha$ for $\alpha\in\set{\x,\y,\z,\1}$ is a Pauli
operator, with $\sigma_\1\equiv\1$; and $\bk{r|\sigma_\alpha|s}$
denotes a matrix element of $\sigma_\alpha$.  This construction
exemplifies how the set of jump operators
$\J_{\t{loss}}^{\t{bare}}\equiv\set{c_{js}}$ that generate particle
loss cannot be constructed from spin operators, which are generally
bilinear in particle creation/annihilation operators.  When working on
the level of a spin model, therefore, we can at best only {\it
  emulate} the effect of particle loss by some indirect means.

To understand the effect of particle loss on collective spin
operators, we first define a single multi-body spin operator
addressing sites $\v j=\set{j_1,j_2,\cdots,j_{\abs{\v j}}}$,
\begin{align}
  \sigma_{\v j\v\alpha} \equiv \prod_{j\in\v j} \sigma_{\alpha_j}^{(j)},
\end{align}
and expand
\begin{align}
  \D\p{\J_{\t{loss}}^{\t{bare}}} \sigma_{\v j\v\alpha}
  &= \sum_{k,s} \p{c_{ks}^\dag \sigma_{\v j\v\alpha} c_{ks}
    - \f12\sp{c_{ks}^\dag c_{ks}, \sigma_{\v j\v\alpha}}_+} \\
  &= \sum_{k\in\v j} \sum_s c_{ks}^\dag \sigma_{\alpha_k}^{(k)} c_{ks}
  \prod_{\substack{j\in\v j\\j\ne k}} \sigma_{\alpha_j}^{(j)}
  + \sum_{k\notin\v j} \sum_s c_{ks}^\dag c_{ks} \sigma_{\v j\v\alpha}
  - \f12 \sum_k \sp{\1^{(k)}, \sigma_{\v j\v\alpha}}_+ \\
  &= \sum_{k\in\v j} \sum_{q,r,s} \bk{q|\sigma_{\alpha_k}|r}
  c_{ks}^\dag c_{kq}^\dag c_{kr} c_{ks}
  \prod_{\substack{j\in\v j\\j\ne k}} \sigma_{\alpha_j}^{(j)}
  - \abs{\v j} \sigma_{\v j\v\alpha}.
\end{align}
In order to have an actual spin model, fermionic statistics or
energetic considerations must forbid multiple occupation of individual
lattice sites.  In that case, the on-site four-point product
$c_{ks}^\dag c_{kq}^\dag c_{kr} c_{ks}=0$ vanishes, and
\begin{align}
  \D\p{\J_{\t{loss}}^{\t{bare}}} \sigma_{\v j\v\alpha}
  = - \abs{\v j} \sigma_{\v j\v\alpha}.
\end{align}
Up to $O(1/N)$ corrections, a collective spin operator $\S_{\v m}$
essentially consists of $\abs{\v m}$-body operators of the form
$\sigma_{\v j\v\alpha}$ with $\abs{\v j}=\abs{\v m}$, which implies
that the dissipator $\D_{\t{loss}}$ defined by
$\D_{\t{loss}}\S_{\v m}=-\abs{\v m}\S_{\v m}$ describes particle loss
with $O(1/N)$ error.  We note that the dissipator $\D_{\t{loss}}$ is
essentially the depolarizing channel,
i.e.~$\D_{\t{loss}}=\D\p{\J_{\t{loss}}}$ for
$\J_{\t{loss}} = \set{s_\alpha^{(j)}}$ with $\alpha\in\set{\x,\y,\z}$
and $j\in\set{1,2,\cdots,N}$.  A direct implementation of
$\D_{\t{loss}}$ with $\D_{\t{loss}}\S_{\v m}=-\abs{\v m}\S_{\v m}$,
however, is much more efficient than evaluating the depolarizing
channel $\D\p{\J_{\t{loss}}}$ with the ingredients in Appendices
\ref{sec:sandwich_single} and \ref{sec:decoherence_single}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robustness of squeezing generated by the twist-and-turn (TNT)
  model}
\label{sec:small_system}
